---
title: "Lab Week 8"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Need to tell R to send stuff back to github
Doesn't automatically perform version control pushing stuff back to github

#####



###Load the packages

```{r}
library(tidyverse)
library(sf)
library(tmap)
library(leaflet)
library(spatstat)
library(maptools)
```


###Column graph of Texas oil spills (look at oils spills csv data)

```{r}
oil_spills <- read_csv("oil_spills.csv")

#create a new data frame and filter oil_spills
df <- oil_spills %>% 
  filter(`Accident State` == "TX" & `Accident Year` < 2017) %>% 
  group_by(`Accident Year`) %>% 
  summarise(Loss = sum(`Net Loss (Barrels)`) ) #create new variable called Loss

#change column names
colnames(df) <- c("Year", "Loss")

ggplot(df, aes(x = Year, y = Loss)) +
  geom_col()
```



Half as much oil spilled in TX land over these 7 years than was spilled in the Exxon-Valdez oil spill


### Leaflet plot of spill locations in TX in 2016

```{r}
df_loc <- oil_spills %>% 
  filter(`Accident State` == "TX" & `Accident Year` == 2016) %>% 
  select(Latitude, Longitude, `Net Loss (Barrels)`)

colnames(df_loc) <- c("latitude", "longitude", "net_loss")

oil_sf <- st_as_sf(df_loc, coords = c("longitude", "latitude"), crs = 4326)

leaflet(oil_sf) %>% 
  addTiles() %>% 
  addMarkers() # only have net loss variable, so don't have to specify

```


## tmap plot with the Texas state shapefile

```{r}
#need to load state boundary files, easy because in working directory

states <- st_read(dsn = ".", layer = "states")
#treat information as a data frame

tex_border <- states %>% 
  filter(STATE_NAME == "Texas") %>% 
  st_transform(4326)

plot(tex_border)

tm_shape(tex_border) +
  tm_polygons() +
  tm_shape(oil_sf) +
  tm_dots(size = 0.3)

```

Are these spatially random? Or are these non-random locations? Does this point pattern follow CRS?


###Convert the data to spatial points patterns
(combination of point data and the bounding window)

```{r}

spill_sp <- as(oil_sf, "Spatial")
spill_ppp <-as(spill_sp, "ppp") #point pattern analysis

tx_sp <- as(tex_border, "Spatial")
tx_owin <- as(tx_sp, "owin") #specify as window for point pattern analysis
#still need to put together

all_ppp <- ppp(spill_ppp$x, spill_ppp$y, window = tx_owin)
#omits points that don't align between window and point pattern

```


###A density plot:
```{r}
plot(density(all_ppp, sigma = 0.4))
```

Density plots are easy to make, but depending on decision making can drastically change the story. For example, changing the sigma value greatly changes the plot. Need to have a meaningful way of deciding the bandwidth (sigma)

```{r}
plot(density(all_ppp, sigma = 0.1))
```




###Quadrant test for spatial evenness

```{r}
oil_qt <- quadrat.test(all_ppp, nx = 5, ny = 5) #pick quadrant size, arbitrary so very dangerous, in this case makes 5 regions and calculates the intensity with chi squared to test if it is even or not
oil_qt #Data are not spatially evenly distributed because p < 0.05
```

Doesn't look like CSR but not obvious visually. See strong clustering in some areas. Also have large blank areas.

Test null hypothesist of CSR (null is that point data follow CSR)

Based on test, reject null hypothesis that data follow CSR. Accept alternative hypothesis that data do not follow CSR (this means that they are not evenly distributed)


```{r}
plot(all_ppp)
#add actual quadrat test info
plot(oil_qt, add = TRUE, cex = 0.4)
```


If have data that is more clustered than CSR, then are nearest neighbors closer together or further apart on average? CLOSER


Graph of distance (r) and proportion of pairs within distance (r), have higher proportion of pairs within shorter distance if more clustered than CSR; if more even than CSR then have fewer point pairs within short distances of each other

Can think about this in two ways:
Single, closest nearest neighbor -> use g function
L&K function -> draw increasing bubbles around each point to see density of neighbors as bubble gets larger to increase more observations

###G-Function for Nearest Neighbor Analysis

```{r}
r <- seq(0,1, by = 0.01)

oil_gfun <- envelope(all_ppp, fun = Gest, r = r, nsim = 100) #give distances over which will calculate proportion of points that we are calling r, run 100 simulations for CSR
#Fast for G function because only looking at nearest neighbor, more computational power needed for L and K function
#theo = theoretical values based on simulation
#runs simulation for CSR data

```
What would data look like if truly CSR? Compare data to that 


Plot data
```{r}
ggplot(oil_gfun, aes(x = r, y = obs)) + 
  geom_line(color = "black") +
  geom_line(aes(x = r, y = theo), color = "red") #compare to CSR
```
Our observed data have a higher proportion of point pairs of nearest neighbors at shorter distances than CSR data -> Data more clustered than modeled CSR data, not truly randomly distributed

r depends on coordinates you are using
WSG84 has degrees, so this would be looking at r being within 1 degree

Instead of using a single nearest neighbor, consider concentrations around a point


###Nearest neighbor using the L-function (Ripley's K standardized)

Similar to using G function

```{r}
r2 <- seq(0,3, by = 0.5) #increment large to save time

oil_lfun <- envelope(all_ppp, fun = Lest, r = r2, nsim = 20, global = TRUE) #because computationally more intensive, running fewer simulations, increases bubble size until covers all observations, global = TRUE to include all observations

```


Create a plot
```{r}
ggplot(oil_lfun, aes(x = r2, y = obs)) +
  geom_line(color = "black") +
  geom_line(aes(x=r2, y = theo), color = "blue")
```


Could calculate maximum nearest neighbor distance to get a starting point, or just play around to get correct size or r2 sequence

Get back to github
















